Evolutionary Quantum Transition (EQUATe): A Novel Solution to Optimization Problems \\

### 2020-09-30 paper_nit.tex
draft4

\title{Evolutionary Quantum Transition (EQT): A Novel Solution to Optimization Problems \\
% {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
% should not be used}
% \thanks{Identify applicable funding agency here. If none, delete this.}
}

@@

\begin{abstract}
Quantum Computing (QC) has often been touted as an esoteric and terrifying field of computing research. However, the possible advantages offered by the inherent quantum fundamentals beseeches extensive additional ventures into this field. Likewise, Evolutionary Computing (EC) offers a multi-pronged approach by deploying several candidates into the search space with constraints guiding the search process. In this paper, we present Evolutionary Quantum Transition to provide a unique solution to a flavour to optimization solving processes using concepts borrowed from QC and EC. The combination leverages compute power to explore the space of candidate solutions by introducing smarter techniques. The solution so obtained provides the optimal transition solution with an automated circuit selection process which can be attributed to the amazing capabilites of QC and EC. 
\end{abstract}

@@

{\scriptsize
\begin{equation}
Optimize \{f\textsubscript{1}(x), f\textsubscript{2}(x), \dots, f\textsubscript{k}(x)\}
\end{equation}
}

@@

EQT utilizes major quantum phenomena such as superposition, quantum entanglment, principle of uncertainity among others \cite{az}, for improving existing search and optimization techniques. Moreover, the space saving capabilities offered by QC are excellent additions to EQT.  The time required for processing a search request decreases exponentially in a QC environment as multiple states are entangled.  
% according to Moore's Law the size of computational units shrinks at an exponential rate as the number of transistors on a chipset increases every year. Even after accounting for physical constraints, a time will come when operations will be conducted on an atomic scale which quantum physics dominates classical physics. As a result, understanding QC and applying them in a virtual landscape to get a sense of their possibilities is a must.

@@

Evolutionary Algorithms (EA) \cite{vik} encompass a wide array of research ideas stemming from general principles of genetics and the theory of evolution. Models are developed to illustrate the behaviour of naturally occuring phenomena, develop these algorithms and test out the application of the corresponding theory. The general process of any EA is depicted in figure \ref{p1}. Borrowing this concept EQT generates an initial population of candidate solutionsand each solution is labelled according to their performance on a set fitness function until a stopping criteria is met. The dominating solution, so obtained, corresponds to the global solution

@@

\subsection{A Novel Ensemble}
As we have seen, each vertical offers it's own advantages and disadvantages when it comes to solving an optimization problem. A trade-off exists between CC and QC when it comes to delivering quality results in a reduced time frame and availability of resources to implement the search process. Meanwhile, EC offers a multi-pronged approach to the search process by applying certain constraints and operators which results in the emergence of the best results with a decreased time complexity. EQT borrows the most promising concepts from QC and EC to provide a new front to solving optimization problems. The general MOOP is a relatively broad set of problems, with each individual problem requiring certain tweaks to the solution search process. In this paper, we narrow down the MOOPs to a specific case i.e, Design a circuit to achieve state $j$ from $i$ with multiple cost considerations. 
% \subsection{Computing Methodologies}
% 
% Most form of computational algorithms in the present day, are executed on a conventional computer. The fundamental notation used for differentiating Classical Computing (CC) and QC \cite{nara} is the basic unit of information. While classical computers use bits 0 and 1, quantum computers use "one of" two computational basis states. The label awarded to these states is "bra-ket" notation i.e, state $\ket{0}$ or $\ket{1}$. Bits are assigned states 0 or 1 deterministically and independently. However, qbuits can exist in a superposition state of $\alpha\textsubscript{0}$$\ket{0}$ + $\alpha\textsubscript{1}$$\ket{1}$, where $\alpha{0}$ and $\alpha{1}$ are complex numbers, such that $|\alpha\textsubscript{0}|\textsuperscript{2}$ + $|\alpha\textsubscript{1}|\textsuperscript{2}$ = 1.
%

@@

Evolved self-organizing computer programs \cite{self} utilize the techniques described in the previous section to solve problems. In this paper the following general problem structure is considered; How to design a circuit to travel from state $\textit{A}$ to state $\textit{B}$? These problem types are called "Oracle Analysis" as they require we to determine some property of a valid quantum gate with no prior information on the intermediate states and what combination of gates results in what output. Some of the key terminologies used in the following texts are explained in Table $\ref{tab:1}$

\begin{table}[!t]
\caption{\textsc{Basic Terminology}}
\label{tab:1}
\centering
{
\begin{tabular}{| c | c |}
\hline
Symbol  & Description \\
\hline
$\psi\textsubscript{i}$ & Current State \\ 
$\psi\textsubscript{o}$ & Desired State \\
$\psi\textsubscript{best}$ & Current Best State \\
$\textit{T}$ & Circuit operator\\
$\textit{F}$ & Fitness function\\
$\textit{M}$ & Number of chromosomes\\
$\textit{G}$ & Number of generations\\
$\textit{sel\_tech}$ & Selection Techniques\\
$\textit{cross\_tech}$ & Crossover Techniques\\
$\textit{mut\_tech}$ & Mutation Techniques\\
$\textit{n}$ & Dimensionality of chromosome\\
$\textit{U(n)}$ & Space of all $\textit{n}$ dimensional matrices\\
$\textit{H\textsubscript{cur}}$ & Current Gate Matrix\\
$\textit{H\textsubscript{cur}\textsuperscript{$\dagger$}}$ & Hermitean Adjoint of $\textit{H\textsubscript{cur}}$\\
$\ket{A\textsubscript{k}}(\psi)$ & Orthonormal basis of the environment Hilbert space\\
$p\textsubscript{k}$ & Input mapped to Hilbert space\\
$q\textsubscript{k}$ & Output mapped to Hilbert space\\
\hline
\end{tabular}
}
\end{table}

@@

\subsection{Modern Approach}

The conventional approach is extremely inefficient and arduous. EQT offers an elegant solution by utilizing the best of both the worlds.

\subsubsection{Applied Quantum Mechanics}
Classicial state transitions cannot fathom superposition and hence will always be fixated on one fixed state. For the D flip flop example, the number of inputs are two and hence the number of output states is $2\textsuperscript{2}$ for each starting state. The 4 possible states can be represented as an one-hot encoded vector with each bit representing the state of the output. For example, the vector shown below represents state 00 for an input state of 0.
$$
\begin{bmatrix} 
1&0&0&0\\
\end{bmatrix}
\quad
$$

Quantum Mechanics is a generalized extension of classical method. The first generalization comes in the form of the temporal values of the elements of the matrix. In QC qubits are utilized instead of bits and as a result the state vectors are variables. Instead of 0 and 1, state vector can be constituted of complex numbers which meet the condition of unitarity. For any matrix $\textit{M}$ to be an unitarity matrix, the only necessary condition is shown in Equation $\ref{eq:1}$ where $M\textsuperscript{$\dagger$}$ is the Hermitean adjoint of $\textit{M}$ and $\textit{I}$ is the identity matrix. 
{\scriptsize
\begin{equation}
\label{eq:1}
M\textsuperscript{$\dagger$}M = MM\textsuperscript{$\dagger$} = I
\end{equation}
}

@@

The process of achieving the target state is discussed in this section. For any test function provided, a set of $\textit{M}$ chromosomes are initialized and based on the input, $\textit{U(n)}$ is laid out. Consider applying set of quantum operators $\textit{Q\textsubscript{$\mu$}}$ on the quantum register of chromosomes for $\mu \in [1,o]$. The output state $\mu$ for an input qubit chromosome $\ket{\psi\textsubscript{i}}$ using operator $\textit{Q\textsubscript{$\mu$}}$, can be determined in a probabilistic manner as shown in Equation $\ref{eq:2}$
{\scriptsize
\begin{equation}
\label{eq:2}
Pr(\mu) = \bra{\psi} Q\textsubscript{$\mu$}\textsuperscript{$\dagger$}Q\textsubscript{$\mu$}\ket{\psi}
\end{equation}
}%
Similarly the probability of a required state $\mu$ is given in Equation $\ref{eq:3}$
{\scriptsize
\begin{equation}
\label{eq:3}
\ket{\mu} = \frac{Q\textsubscript{$\mu$}\ket{\psi}}{\sqrt{Pr(\mu)}}
\end{equation}
}
Since the states can be unitary i.e $\textit{H\textsubscript{cur}} \cdot$ $\textit{H\textsubscript{cur}\textsuperscript{$\dagger$}} = I$, elements of $\textit{U(n)}$ will have a definite form. For example, any element of $\textit{U(2)}$, say $\textit{T}$, will have the following format for complex inputs $z\textsubscript{1}$ and $z\textsubscript{2}$ seperated by an angle $\theta$:
$$
\begin{bmatrix} 
{z\textsubscript{1}}&{z\textsubscript{2}}\\
{-e\textsuperscript{i$\theta$}z\textsubscript{1}\textsuperscript{*}}&{-e\textsuperscript{i$\theta$}z\textsubscript{2}\textsuperscript{*}}\\
\end{bmatrix}
\quad
$$

@@

\begin{algorithm}[!b]
\footnotesize
\caption{Evolutionary Quantum Algorithm}
\label{alg:eq}
\begin{algorithmic}[1]
\STATE \textbf{Input:} $\psi\textsubscript{i}$, $\psi\textsubscript{o}$, $\textit{F}$, $\textit{M}$, $\textit{G}$, $\textit{sel\_tech}$, $\textit{cross\_tech}$ and $\textit{mut\_tech}$
\STATE Identify $\textit{n}$ as dimension of input
\STATE Create a grid encompassing all combinations of $\textit{sel\_tech}$, $\textit{cross\_tech}$ and $\textit{mut\_tech}$
\FOR{Each generation $\textit{G}$}
\STATE Evaluate the chromosomes against $\textit{F}$
\STATE Call Algorithm $\ref{alg:gwo}$ with the grid, $\psi\textsubscript{i}$ and $\psi\textsubscript{o}$ 
\STATE Obtain returned $\psi\textsubscript{best}$
\ENDFOR
\STATE Select the optimal $\psi\textsubscript{best}$
\end{algorithmic}
\end{algorithm}

\subsection{Applied Evolutionary Operators}

@@

In this paper, we follow a genetic modified Grey Wolf Optimization (GWO) algorithm \cite{gwo} to obtain and update candidate solutions. GWO follows a hierarchical structure of candidate solution which imitates the behaviour of Grey Wolves as observed in nature. The structure of $\alpha$, $\beta$, $\delta$ and $\omega$ labelled wolves. The $\alpha$ wolves are responsible for finding the global optimum whilst being guided by the $\beta$ wolves. The $\delta$ and $\omega$ provide secondary assistance to the leaders and hence maitain the dominant structure. After each search iteration, genetic operators like selection, crossover and mutation are applied to every candidate in the structure. The entire process is discribed in Algorithm $\ref{alg:gwo}$

\begin{algorithm}[t]
\footnotesize
\caption{GWO pseudocode}
\label{alg:gwo}
\begin{algorithmic}[1]
\STATE \textbf{Input:} Population size $\textit{n}$, random vectors $\vec{r\textsubscript{1}}$, $\vec{r\textsubscript{2}} \in$ [0,1], initial prey location $\vec{D}$, number of iterations \textit{I}, fitness function $\textit{f}$, coefficient vectors $\vec{A}$, $\vec{C}$ and $\vec{a}$
\STATE Set $\textit{t}$ = 0
\FOR{i $\in$ [1,n]}    
\STATE Generate wolf pack population $\textit{X\textsubscript{i}(t)}$ at instance $\textit{t}$
\STATE Evaluate each individual against the fitness function
\ENDFOR
\STATE Assign $\alpha, \beta, \delta$ titles to the top three solutions
\STATE Evaluate $\vec{D} = |\vec{C} \cdot \vec{X\textsubscript{p}}(t) - \vec{X}(t)|$
\FOR{\textit{i} in \textit{I}}
\FOR{Each individual in $\textit{n}$}  
\STATE $\vec{X\textsubscript{1}} = \vec{X\textsubscript{$\alpha$}} - \vec{A\textsubscript{1}} \cdot  \vec{D\textsubscript{$\alpha$}}$
\STATE $\vec{X\textsubscript{2}} = \vec{X\textsubscript{$\beta$}} - \vec{A\textsubscript{2}} \cdot  \vec{D\textsubscript{$\beta$}}$
\STATE $\vec{X\textsubscript{3}} = \vec{X\textsubscript{$\delta$}} - \vec{A\textsubscript{3}} \cdot  \vec{D\textsubscript{$\delta$}}$
\STATE Evaluate $\vec{X}(t+1) = \frac{\vec{X\textsubscript{1}} + \vec{X\textsubscript{2}} + \vec{X\textsubscript{3}}}{3}$
\ENDFOR
\STATE Update the coefficient vectors $\vec{A}$ and $\vec{C}$
\STATE $\vec{A} = 2\vec{a} \cdot \vec{r\textsubscript{1}} - \vec{a}$
\STATE $\vec{C} = 2\vec{r\textsubscript{2}}$
\STATE Linearly decrease $\vec{a}$ from 2 to 0
% \STATE Update $\vec{X\textsubscript{$\alpha$}}, \vec{X\textsubscript{$\beta$}}, \vec{X\textsubscript{$\delta$}}$
\STATE Perform Crossover, Selection and Mutation
\STATE Increment $\textit{t}$
\ENDFOR
\STATE $\vec{X\textsubscript{$\alpha$}}$ corresponds to the global optimum. 
\end{algorithmic}
\end{algorithm}

The genetic operators for a sample candidate in the form of a genomic are explained here following which the process is depicted in Figure :
\begin{itemize}
\item Selection: Given a population of potential solutions, selection involves grading these solutions against a fitness function. The selected solutions are cast into the crossover stage.
\item Crossover: From the selected set of solutions, randomly two solutions are selected and the next generation chromosome is created which shares the best attributes of it's parents.
\item Mutation: The selected solutions undergo random changes in their composition to promote diversity and increase their changes of converging to the global optimum.
\end{itemize}

\begin{figure}[!t]
\centering\includegraphics[width=7cm]{p2.png}
\caption{Genetic Operators}
\label{p3}
\end{figure}

As an example consider, initial state as [1 1 1 0 0 0 0 1] and the target state as [1 0 1 0 0 1 0 0]. Applying the concepts discussed previously, a quantum circuit gate is created for this transition and the associated matrix is given by 
$$
\begin{bmatrix}
\begin{bmatrix}
-0.14112954-0.16700839j& 0.07983181-0.29873381j\\
-0.25803742+0.38299128j& 0.22384403+0.01717024j\\
0.06272906+0.25203229j& -0.17352533+0.18676704j\\
-0.46040214-0.12675132j& -0.11769043+0.46751876j\\
\end{bmatrix}\\
\begin{bmatrix}
0.31172643+0.14024689j& -0.11225685-0.02087697j\\
0.16743502+0.29588733j&  0.18771276+0.01118339j\\
0.57520561+0.33055889j&  0.13856821-0.24062621j\\
0.15502007-0.07663589j& -0.38171798-0.16226323j\\
\end{bmatrix}\\
\begin{bmatrix}
0.23308965-0.15943424j& 0.26838968-0.17899447j\\
-0.19340125+0.42085285j&-0.1072532 +0.1531806j\\
0.04524549-0.44413067j& 0.15183147-0.24693186j\\
-0.2982411 +0.05650701j&  0.18590532-0.3958033j\\
\end{bmatrix}\\
\begin{bmatrix}
-0.07005404+0.04745812j& -0.27252377-0.17790739j\\
-0.2187864 +0.24156256j& -0.26294328+0.3396825j\\
0.10360916-0.35043418j&  0.12404486+0.42213862j\\
0.44304366-0.02873782j& -0.22363463+0.14806583j\\
\end{bmatrix}\\
\begin{bmatrix}
0.37175985+0.28751052j &-0.07056257-0.1013915j\\  
-0.35527493+0.07918063j & 0.48963077-0.31625964j\\
-0.20088489-0.20871619j &-0.19892169-0.09136702j\\
0.29164278+0.21288127j  &0.08750707+0.14750696j\\
\end{bmatrix}\\
\begin{bmatrix}
-0.38421248-0.04067692j& 0.16637992+0.61802823j\\ 
-0.17373866+0.27097997j& 0.18105618-0.29023612j \\
0.02505396-0.15353303j& 0.11080511+0.16151108j\\
-0.02153527+0.158645j& -0.32556212-0.16206476j\\
\end{bmatrix}\\
\begin{bmatrix}
0.19075161-0.30210472j& -0.09189932-0.2130557j\\  
-0.13117622-0.29864423j&  0.01177846-0.25807943j\\
-0.11828712+0.02111738j&  0.687924  +0.04122678j\\
-0.12997132+0.21518361j& -0.28512878+0.10722509j\\
\end{bmatrix}\\
\begin{bmatrix}
0.50059262+0.00708867j& -0.01602864+0.44984401j\\  
0.01398646+0.08782445j& -0.25293655+0.32748978j\\
-0.13217757+0.13982117j& -0.11740198+0.07269765j\\
-0.1949218 +0.44227633j& -0.09510568+0.26242716j\\
\end{bmatrix}\\

\end{bmatrix}
$$
\subsection{Key Points}
Some of the important points governing the flow of the algorithm are mentioned below:

@@

\item The most important characteristic of this endeavour is the ability of both the sub-algorithm to retain information. For the evolutionary part, it is the lower swarm particles who memorize the pitfalls. Information retention in QC is guarenteed by the No-Hiding Theorem. The No-Hiding Theorem in an enlarged Hilbert space is given in Equation $\ref{eq:4}$. Due to the correlation between subsystems, quantum information cannot be lost. 
{\scriptsize
\begin{equation}
\label{eq:4}
\begin{split}
\ket{\mu} \otimes \ket{A} & \rightarrow \sum_{k}^{}\sqrt{p\textsubscript{k}}\ket{k} \otimes \ket{A\textsubscript{k}(\psi)}\\
& = \sum_{k}^{}\sqrt{p\textsubscript{k}}\ket{k} \otimes (\ket{q\textsubscript{k}} \otimes \ket{psi} \otimes 0)
\end{split}
\end{equation}
}%
\item The unused dimension of the Hilbert space are filled by $\otimes 0$ vectors. These unused dimensions can be utilized to completely hide a subsystem and hence delete the information.
\end{itemize}

>>>>

\title{Evolutionary Quantum Transition (EQUATe): A Novel Solution to Optimization Problems \\
% {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
% should not be used}
% \thanks{Identify applicable funding agency here. If none, delete this.}
}

@@

\begin{abstract}
Quantum Computing (QC) has often been touted as an esoteric and terrifying field of computing research. However, the possible advantages offered by the inherent quantum fundamentals beseeches extensive additional ventures into this field. Likewise, Evolutionary Computing (EC) offers a multi-pronged approach by deploying several candidates into the search space with constraints guiding the search process. In this paper, we present Evolutionary Quantum Transition (EQUATe) to provide a unique solution to solving Multi-Optimization Objective Problems (MOOPs) using concepts borrowed from QC and EC. The combination leverages compute power to explore the space of candidate solutions by introducing smarter techniques. The solution so obtained provides the optimal transition solution with an automated circuit selection process which can be attributed to the amazing capabilites of QC and EC. 
\end{abstract}

@@

\begin{equation}
Optimize \{f\textsubscript{1}(x), f\textsubscript{2}(x), \dots, f\textsubscript{k}(x)\}
\end{equation}

@@

QC utilizes major quantum phenomena such as superposition, quantum entanglment, principle of uncertainity among others \cite{az}, for improving existing search and optimization techniques. The time required for processing a search request decreases exponentially in a QC environment as multiple states are entangled.  
% according to Moore's Law the size of computational units shrinks at an exponential rate as the number of transistors on a chipset increases every year. Even after accounting for physical constraints, a time will come when operations will be conducted on an atomic scale which quantum physics dominates classical physics. As a result, understanding QC and applying them in a virtual landscape to get a sense of their possibilities is a must.

@@

Evolutionary Algorithms (EA) \cite{vik} encompass a wide array of research ideas stemming from general principles of genetics and the theory of evolution. Models are developed to illustrate the behaviour of naturally occuring phenomena, develop these algorithms and test out the application of the corresponding theory. The general process of any EA is depicted in figure \ref{p1}.

@@

\subsection{A Novel Ensemble}
As we have seen, each vertical offers it's own advantages and disadvantages when it comes to solving an optimization problem. A trade-off exists between CC and QC when it comes to delivering quality results in a reduced time frame and availability of resources to implement the search process. Meanwhile, EC offers a multi-pronged approach to the search process by applying certain constraints and operators which results in the emergence of the best results with a decreased time complexity. EQUATe borrows the most promising concepts from QC and EC to provide a new front to solving optimization problems. The space saving capabilities offered by QC are excellent additions to EQUATe. Likewise the evolutionary model offered by EC boost the search process. The general MOOP is a relatively broad set of problems, with each individual problem requiring certain tweaks to the solution search process. In this paper, we narrow down the MOOPs to a specific case i.e, Design a circuit to achieve state $j$ from $i$ with multiple cost considerations. 
% \subsection{Computing Methodologies}
% 
% Most form of computational algorithms in the present day, are executed on a conventional computer. The fundamental notation used for differentiating Classical Computing (CC) and QC \cite{nara} is the basic unit of information. While classical computers use bits 0 and 1, quantum computers use "one of" two computational basis states. The label awarded to these states is "bra-ket" notation i.e, state $\ket{0}$ or $\ket{1}$. Bits are assigned states 0 or 1 deterministically and independently. However, qbuits can exist in a superposition state of $\alpha\textsubscript{0}$$\ket{0}$ + $\alpha\textsubscript{1}$$\ket{1}$, where $\alpha{0}$ and $\alpha{1}$ are complex numbers, such that $|\alpha\textsubscript{0}|\textsuperscript{2}$ + $|\alpha\textsubscript{1}|\textsuperscript{2}$ = 1.
%

@@

Evolved self-organizing computer programs \cite{self} utilize the techniques described in the previous section to solve problems. In this paper the following general problem structure is considered for demonstrating the capabilities of EQUATe; How to design a circuit to travel from state $\textit{A}$ to state $\textit{B}$? These problem types are called "Oracle Analysis" as they require we to determine some property of a valid quantum gate with no prior information on the intermediate states and what combination of gates results in what output. Some of the key terminologies used in the following texts are explained in Table $\ref{tab:2}$

\begin{table}[!t]
\caption{\textsc{Basic Terminology}}
\label{tab:2}
\centering
{
\begin{tabular}{| c | c |}
\hline
Symbol  & Description \\
\hline
$\psi\textsubscript{i}$ & Current State \\ 
$\psi\textsubscript{o}$ & Desired State \\
$\psi\textsubscript{best}$ & Current Best State \\
$\textit{T}$ & Circuit operator\\
$\textit{F}$ & Fitness function\\
$\textit{M}$ & Number of chromosomes\\
$\textit{G}$ & Number of generations\\
$\textit{sel\_tech}$ & Selection Techniques\\
$\textit{cross\_tech}$ & Crossover Techniques\\
$\textit{mut\_tech}$ & Mutation Techniques\\
$\textit{n}$ & Dimensionality of chromosome\\
$\textit{U(n)}$ & Space of all $\textit{n}$ dimensional matrices\\
$\textit{H\textsubscript{cur}}$ & Current Gate Matrix\\
$\textit{H\textsubscript{cur}\textsuperscript{$\dagger$}}$ & Hermitean Adjoint of $\textit{H\textsubscript{cur}}$\\
$\ket{A\textsubscript{k}}(\psi)$ & Orthonormal basis of the environment Hilbert space\\
$p\textsubscript{k}$ & Input mapped to Hilbert space\\
$q\textsubscript{k}$ & Output mapped to Hilbert space\\
\hline
\end{tabular}
}
\end{table}

@@

\subsection{EQUATe Approach}

The conventional approach is extremely inefficient and arduous. EQUATe offers an elegant solution by utilizing the best of both the worlds.

\subsubsection{Applied Quantum Mechanics for EQUATe}
Classicial state transitions cannot fathom superposition and hence will always be fixated on one fixed state. For the D flip flop example, the number of inputs are two and hence the number of output states is $2\textsuperscript{2}$ for each starting state. The 4 possible states can be represented as an one-hot encoded vector with each bit representing the state of the output. For example, the vector shown below represents state 00 for an input state of 0.
$$
\begin{bmatrix} 
1&0&0&0\\
\end{bmatrix}
\quad
$$

Quantum Mechanics is a generalized extension of classical method. The first generalization comes in the form of the temporal values of the elements of the matrix. In QC qubits are utilized instead of bits and as a result the state vectors are variables. Instead of 0 and 1, state vector can be constituted of complex numbers which meet the condition of unitarity. For any matrix $\textit{M}$ to be an unitarity matrix, the only necessary condition is shown in Equation $\ref{eq:1}$ where $M\textsuperscript{$\dagger$}$ is the Hermitean adjoint of $\textit{M}$ and $\textit{I}$ is the identity matrix.

\begin{equation}
\label{eq:1}
M\textsuperscript{$\dagger$}M = MM\textsuperscript{$\dagger$} = I
\end{equation}

@@

The process of achieving the target state is discussed in this section. For any test function provided, a set of $\textit{M}$ chromosomes are initialized and based on the input, $\textit{U(n)}$ is laid out. Consider applying set of quantum operators $\textit{Q\textsubscript{$\mu$}}$ on the quantum register of chromosomes for $\mu \in [1,o]$. The output state $\mu$ for an input qubit chromosome $\ket{\psi\textsubscript{i}}$ using operator $\textit{Q\textsubscript{$\mu$}}$, can be determined in a probabilistic manner as shown in Equation $\ref{eq:2}$

\begin{equation}
\label{eq:2}
Pr(\mu) = \bra{\psi} Q\textsubscript{$\mu$}\textsuperscript{$\dagger$}Q\textsubscript{$\mu$}\ket{\psi}
\end{equation}

Similarly the probability of a required state $\mu$ is given in Equation $\ref{eq:3}$

\begin{equation}
\label{eq:3}
\ket{\mu} = \frac{Q\textsubscript{$\mu$}\ket{\psi}}{\sqrt{Pr(\mu)}}
\end{equation}

Since the states can be unitary i.e $\textit{H\textsubscript{cur}} \cdot$ $\textit{H\textsubscript{cur}\textsuperscript{$\dagger$}} = I$, elements of $\textit{U(n)}$ will have a definite form. For example, any element of $\textit{U(2)}$, say $\textit{T}$, will have the following format for complex inputs $z\textsubscript{1}$ and $z\textsubscript{2}$ seperated by an angle $\theta$:
$$
\begin{bmatrix} 
{z\textsubscript{1}}&{z\textsubscript{2}}\\
{-e\textsuperscript{i$\theta$}z\textsubscript{1}\textsuperscript{*}}&{-e\textsuperscript{i$\theta$}z\textsubscript{2}\textsuperscript{*}}\\
\end{bmatrix}
\quad
$$

@@

\begin{algorithm}[!b]
\footnotesize
\caption{EQUATe}
\label{alg:eq}
\begin{algorithmic}[1]
\STATE \textbf{Input:} $\psi\textsubscript{i}$, $\psi\textsubscript{o}$, $\textit{F}$, $\textit{M}$, $\textit{G}$, $\textit{sel\_tech}$, $\textit{cross\_tech}$ and $\textit{mut\_tech}$
\STATE Identify $\textit{n}$ as dimension of input
\STATE Create a grid encompassing all combinations of $\textit{sel\_tech}$, $\textit{cross\_tech}$ and $\textit{mut\_tech}$
\FOR{Each generation $\textit{G}$}
\STATE Evaluate the chromosomes against $\textit{F}$
\STATE Call Algorithm $\ref{alg:gwo}$ with the grid, $\psi\textsubscript{i}$ and $\psi\textsubscript{o}$ 
\STATE Obtain returned $\psi\textsubscript{best}$
\ENDFOR
\STATE Select the optimal $\psi\textsubscript{best}$
\end{algorithmic}
\end{algorithm}

\subsection{Applied Evolutionary Operators for EQUATe}

@@

In this paper, we follow a genetically modified Grey Wolf Optimization (GWO) algorithm \cite{gwo} to obtain and update candidate solutions. GWO follows a hierarchical structure of candidate solution which imitates the behaviour of Grey Wolves as observed in nature. The structure of $\alpha$, $\beta$, $\delta$ and $\omega$ labelled wolves. The $\alpha$ wolves are responsible for finding the global optimum whilst being guided by the $\beta$ wolves. The $\delta$ and $\omega$ provide secondary assistance to the leaders and hence maitain the dominant structure. After each search iteration, genetic operators like selection, crossover and mutation are applied to every candidate in the structure. The entire process is discribed in Algorithm $\ref{alg:gwo}$

\begin{algorithm}[t]
\footnotesize
\caption{GWO pseudocode}
\label{alg:gwo}
\begin{algorithmic}[1]
\STATE \textbf{Input:} Population size $\textit{n}$, random vectors $\vec{r\textsubscript{1}}$, $\vec{r\textsubscript{2}} \in$ [0,1], initial prey location $\vec{D}$, number of iterations \textit{I}, fitness function $\textit{f}$, coefficient vectors $\vec{A}$, $\vec{C}$ and $\vec{a}$
\STATE Set $\textit{t}$ = 0
\FOR{i $\in$ [1,n]}    
\STATE Generate wolf pack population $\textit{X\textsubscript{i}(t)}$ at instance $\textit{t}$
\STATE Evaluate each individual against the fitness function
\ENDFOR
\STATE Assign $\alpha, \beta, \delta$ titles to the top three solutions
\STATE Evaluate $\vec{D} = |\vec{C} \cdot \vec{X\textsubscript{p}}(t) - \vec{X}(t)|$
\FOR{\textit{i} in \textit{I}}
\FOR{Each individual in $\textit{n}$}  
\STATE $\vec{X\textsubscript{1}} = \vec{X\textsubscript{$\alpha$}} - \vec{A\textsubscript{1}} \cdot  \vec{D\textsubscript{$\alpha$}}$
\STATE $\vec{X\textsubscript{2}} = \vec{X\textsubscript{$\beta$}} - \vec{A\textsubscript{2}} \cdot  \vec{D\textsubscript{$\beta$}}$
\STATE $\vec{X\textsubscript{3}} = \vec{X\textsubscript{$\delta$}} - \vec{A\textsubscript{3}} \cdot  \vec{D\textsubscript{$\delta$}}$
\STATE Evaluate $\vec{X}(t+1) = \frac{\vec{X\textsubscript{1}} + \vec{X\textsubscript{2}} + \vec{X\textsubscript{3}}}{3}$
\ENDFOR
\STATE Update the coefficient vectors $\vec{A}$ and $\vec{C}$
\STATE $\vec{A} = 2\vec{a} \cdot \vec{r\textsubscript{1}} - \vec{a}$
\STATE $\vec{C} = 2\vec{r\textsubscript{2}}$
\STATE Linearly decrease $\vec{a}$ from 2 to 0
% \STATE Update $\vec{X\textsubscript{$\alpha$}}, \vec{X\textsubscript{$\beta$}}, \vec{X\textsubscript{$\delta$}}$
\STATE Perform Crossover, Selection and Mutation
\STATE Increment $\textit{t}$
\ENDFOR
\STATE $\vec{X\textsubscript{$\alpha$}}$ corresponds to the global optimum. 
\end{algorithmic}
\end{algorithm}
The genetic operators for a sample candidate in the form of genomes are explained here following which the process is depicted in Figure $\ref{p3}$:
\begin{itemize}
\item Selection: Given a population of potential solutions, selection involves grading these solutions against a fitness function. The selected solutions are cast into the crossover stage.
\item Crossover: From the selected set of solutions, randomly two solutions are selected and the next generation chromosome is created which shares the best attributes of it's parents.
\item Mutation: The selected solutions undergo random changes in their composition to promote diversity and increase their changes of converging to the global optimum.
\end{itemize}
\begin{figure}[!t]
\centering\includegraphics[width=7cm]{p2.png}
\caption{Genetic Operators}
\label{p3}
As an example consider, initial state as [1 1 1 0 0 0 0 1] and the target state as [1 0 1 0 0 1 0 0]. Applying the concepts discussed previously, a quantum circuit gate is created for this transition and the associated matrix is given by 
{\scriptsize
$$
\begin{bmatrix}
\begin{bmatrix}
-0.14112954-0.16700839j& 0.07983181-0.29873381j\\
-0.25803742+0.38299128j& 0.22384403+0.01717024j\\
0.06272906+0.25203229j& -0.17352533+0.18676704j\\
-0.46040214-0.12675132j& -0.11769043+0.46751876j\\
\end{bmatrix}\\
\begin{bmatrix}
0.31172643+0.14024689j& -0.11225685-0.02087697j\\
0.16743502+0.29588733j&  0.18771276+0.01118339j\\
0.57520561+0.33055889j&  0.13856821-0.24062621j\\
0.15502007-0.07663589j& -0.38171798-0.16226323j\\
\end{bmatrix}\\
\begin{bmatrix}
0.23308965-0.15943424j& 0.26838968-0.17899447j\\
-0.19340125+0.42085285j&-0.1072532 +0.1531806j\\
0.04524549-0.44413067j& 0.15183147-0.24693186j\\
-0.2982411 +0.05650701j&  0.18590532-0.3958033j\\
\end{bmatrix}\\
\begin{bmatrix}
-0.07005404+0.04745812j& -0.27252377-0.17790739j\\
-0.2187864 +0.24156256j& -0.26294328+0.3396825j\\
0.10360916-0.35043418j&  0.12404486+0.42213862j\\
0.44304366-0.02873782j& -0.22363463+0.14806583j\\
\end{bmatrix}\\
\begin{bmatrix}
0.37175985+0.28751052j &-0.07056257-0.1013915j\\  
-0.35527493+0.07918063j & 0.48963077-0.31625964j\\
-0.20088489-0.20871619j &-0.19892169-0.09136702j\\
0.29164278+0.21288127j  &0.08750707+0.14750696j\\
\end{bmatrix}\\
\begin{bmatrix}
-0.38421248-0.04067692j& 0.16637992+0.61802823j\\ 
-0.17373866+0.27097997j& 0.18105618-0.29023612j \\
0.02505396-0.15353303j& 0.11080511+0.16151108j\\
-0.02153527+0.158645j& -0.32556212-0.16206476j\\
\end{bmatrix}\\
\begin{bmatrix}
0.19075161-0.30210472j& -0.09189932-0.2130557j\\  
-0.13117622-0.29864423j&  0.01177846-0.25807943j\\
-0.11828712+0.02111738j&  0.687924  +0.04122678j\\
-0.12997132+0.21518361j& -0.28512878+0.10722509j\\
\end{bmatrix}\\
\begin{bmatrix}
0.50059262+0.00708867j& -0.01602864+0.44984401j\\  
0.01398646+0.08782445j& -0.25293655+0.32748978j\\
-0.13217757+0.13982117j& -0.11740198+0.07269765j\\
-0.1949218 +0.44227633j& -0.09510568+0.26242716j\\
\end{bmatrix}\\
\end{bmatrix}
$$
}
\subsection{Key Points}
Some of the important points governing the flow of the algorithm are mentioned below:

@@

\item The most important characteristic of this endeavour is the ability of both the sub-algorithm to retain information. For the evolutionary part, it is the lower swarm particles who memorize the pitfalls. Information retention in QC is guarenteed by the No-Hiding Theorem. The No-Hiding Theorem in an enlarged Hilbert space is given in Equation $\ref{eq:4}$. Due to the correlation between subsystems, quantum information cannot be lost.

\begin{equation}
\label{eq:4}
\begin{split}
\ket{\mu} \otimes \ket{A} & \rightarrow \sum_{k}^{}\sqrt{p\textsubscript{k}}\ket{k} \otimes \ket{A\textsubscript{k}(\psi)}\\
& = \sum_{k}^{}\sqrt{p\textsubscript{k}}\ket{k} \otimes (\ket{q\textsubscript{k}} \otimes \ket{psi} \otimes 0)
\end{split}
\end{equation}

\item The unused dimension of the Hilbert space are filled by $\otimes 0$ vectors. These unused dimensions can be utilized to completely hide a subsystem and hence delete the information.
\end{itemize}
